Action Groupoids and Fractional Types

### 2016-10-12 pibackground.tex
move forward with LIPIcs format

\noindent Both classes of programs, 1-combinators $c$, and
2-combinators~$\alpha$, denote \emph{equivalences} in the HoTT
sense. The elements $c$ of 1-combinators denote type isomorphisms. The
elements $\alpha$ of 2-combinators denote equivalences between these
type isomorphisms which are sound and complete 1-combinator program
transformations. Using the 1-combinators, it is possible to write any
reversible boolean function and hence encode arbitrary boolean
functions by a technique that goes back to \citet{Toffoli:1980}. The
2-combinators provide a layer of programs that computes
semantic-preserving transformations of 1-combinators. As a small
example, let us abbreviate $\ot \oplus \ot$ as the type $\mathbb{2}$
of booleans and examine two possible implementations of boolean
negation. The first directly uses the primitive combinator
$\swapp : \tau_1 \oplus \tau_2 \iso \tau_2 \oplus \tau_1$ to swap the
two values of type $\mathbb{2}$; the second use three consecutive
copies of $\swapp$ to achieve the same effect:
\[\begin{array}{rcl}
\mathsf{not_1} &=& \swapp \\
\mathsf{not_2} &=& (\swapp \odot \swapp) \odot \swapp 
\end{array}\]
We can write a 2-combinator whose \emph{type} is $\mathsf{not_2}
\isotwo \mathsf{not_1}$:
\[
(\idldl ~\respstwo~ \idisotwo)~\transtwo~\idldl
\]
which not only shows the equivalence of the two implementations of
negation but also shows \emph{how} to transform one to the other.

>>>>

\noindent Both classes of programs, 1-combinators $c$, and
2-combinators~$\alpha$, denote \emph{equivalences} in the HoTT
sense. The elements $c$ of 1-combinators denote type isomorphisms. The
elements $\alpha$ of 2-combinators denote equivalences between these
type isomorphisms which are sound and complete 1-combinator program
transformations. Using the 1-combinators, it is possible to write any
reversible boolean function and hence encode arbitrary boolean
functions by a technique that goes back to \cite{Toffoli:1980}. The
2-combinators provide a layer of programs that computes
semantics-preserving transformations of 1-combinators. As a small
example, let us abbreviate $\ot \oplus \ot$ as the type $\mathbb{2}$
of booleans and examine two possible implementations of boolean
negation. The first directly uses the primitive combinator
$\swapp : \tau_1 \oplus \tau_2 \iso \tau_2 \oplus \tau_1$ to swap the
two values of type $\mathbb{2}$; the second use three consecutive
copies of $\swapp$ to achieve the same effect:
\[\begin{array}{rcl}
\mathsf{not_1} &=& \swapp \\
\mathsf{not_2} &=& (\swapp \odot \swapp) \odot \swapp
\end{array}\]
We can write a 2-combinator whose \emph{type} is $\mathsf{not_2}
\isotwo \mathsf{not_1}$:
\[
(\idldl ~\respstwo~ \idisotwo)~\transtwo~\idldl
\]
which not only shows the equivalence of the two implementations of
negation but also shows \emph{how} to transform one to the other.
